# Reproducible Research: Peer Assessment 1


## Getting the data
The first thing is to check if the zip file has already been downloaded. If not, download it.
```{r}
url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
zip.file <- "activity.zip"
file <- "activity.csv"

if(!file.exists(zip.file)) {
    download.file(url, destfile = zip.file)
    }
```

If the data is not unzipped yet, extract it to the current directory.
```{r}
if(!file.exists(file)) {
    unzip(zip.file)
    }
```

## Loading and preprocessing the data

We read the data:
```{r}
data <- read.csv("activity.csv")
```

Is the data in its correct format?
```{r}
str(data)
```

The date variable is read as a factor, not a date. Let's correct that.
```{r}
data$date <- as.Date(as.character(data$date))
str(data)
```

Ok, that's better. Let's do the homework now.

## What is mean total number of steps taken per day?
Let's first calculate the total number of steps per day.
```{r}
library(plyr)
data.day <- ddply(data, .(date), summarize, sum=sum(steps))
data.day <- tapply(data$steps, data$date, sum)
data.day <- as.data.frame(data.day)
```



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
